COVID-19 dataset strength

### 2020-05-19 bbbq_article.tex
Need to assess data set strength from script
Fix merge conflict (?) at Kolmogorov-Smirnoff test
EOD

\begin{itemize}
	\item Genome size: $\approx 3 \cdot 10^4$ bases
	\item Evolution rate: $\approx 10^{-3}$ per nucleotide per year (from \cite{bar2020sars})
	\item Evolution rate per genome: $\approx 26$ per year (GISAID)
	\item Evolution rate per nucleotide: $0.00088$ per nucleotide per year as $\frac{26}{\cdot 10^4}$ per nucleotide year
	\item Mutation rate: $\approx 10^{-6}$ per nucleotide per cycle
	\item Proteome size: 9701 (from reference proteome)
	\item Size of proteome compared to genome: $\frac{9701}{3 \cdot 10^4} \approx \frac{1}{3}$ \richel{no/little non-coding DNA indeed}
	\item Fraction of mutations that change AA: $\frac{20}{64}$ 
	\item Evolution rate: $\approx 0.00092$ per amino acid per year (as $\frac{1}{3} \cdot \frac{20}{64} \cdot 0.00088$)
	\item TMHs per proteome: 21 (from TMHMM)
	\item AAs per TMHs: 23 (from Bianchi et al.)
	\item AAs per proteome that is TMH: $21 \cdot 23 = 483$
	\item Mutations in TMH per proteome year: $\approx 0.444$ (as $483 \cdot 0.00092$)
	\item Number of sequenced proteomes: 5600 (GISAID, genomes)
	\item Pandemic duration: $\approx \frac{1}{3}$ year
	\item Expected number of observed TMH mutations: $\approx 800$ (as $0.444 \cdot 5600 \cdot \frac{1}{3}$)
	\item Chance that a mutation may lose/acquire epitope binding: $\frac{1}{100}$ \richel{Just a guess, based on weak selection}
	\item Expected number of observed TMH mutations that cause a transition: $\approx 8$ (as $800 \cdot \frac{1}{100}$)
\end{itemize}

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{tmhs/tmhs.png}
  \caption{
    Location of amino acids of reference COVID-19 proteome.
    Note that ORF1a results in multiple proteins, 
    as shown in figure \ref{fig:covid_genome_and_proteome}.
    Legend: i = inside (cytosolic side), m = membrane, o = outside (extracellular side)
    \richel{This is real data}
  }
  \label{fig:covid_topology}
\end{figure}

@@

\end{document}
new research. 
      This table be deleted in the future.
    }
  }
  \label{table:mhc2_haplotypes}
\end{table}

>>>>

\begin{itemize}
	\item Genome size: $\approx 3 \cdot 10^4$ bases
	\item Evolution rate: $\approx 10^{-3}$ per nucleotide per year (from \cite{bar2020sars})
	\item Evolution rate per genome: $\approx 26$ per year (GISAID)
	\item Evolution rate per nucleotide: $0.00088$ per nucleotide per year as $\frac{26}{\cdot 10^4}$ per nucleotide year
	\item Mutation rate: $\approx 10^{-6}$ per nucleotide per cycle
	\item Proteome size: 9701 (from reference proteome)
	\item Size of proteome compared to genome: $\frac{9701}{3 \cdot 10^4} \approx \frac{1}{3}$ \richel{no/little non-coding DNA indeed}
	\item Fraction of mutations that change AA: $\frac{20}{64}$ 
	\item Evolution rate: $\approx 0.00092$ per amino acid per year (as $\frac{1}{3} \cdot \frac{20}{64} \cdot 0.00088$)
	\item TMHs per proteome: 26 (from table \ref{table:covid_tmhs})
	\item AAs per TMHs: 23 (from Bianchi et al.)
	\item AAs per proteome that is TMH: $21 \cdot 23 = 483$
	\item Mutations in TMH per proteome year: $\approx 0.444$ (as $483 \cdot 0.00092$)
	\item Number of sequenced proteomes: 5600 (GISAID, genomes)
	\item Pandemic duration: $\approx \frac{1}{3}$ year
	\item Expected number of observed TMH mutations: $\approx 800$ (as $0.444 \cdot 5600 \cdot \frac{1}{3}$)
	\item Chance that a mutation may lose/acquire epitope binding: $\frac{1}{100}$ \richel{Just a guess, based on weak selection}
	\item Expected number of observed TMH mutations that cause a transition: $\approx 8$ (as $800 \cdot \frac{1}{100}$)
\end{itemize}

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{tmhs/tmhs.png}
  \caption{
    Location of amino acids of reference COVID-19 
    proteome (Uniprot ID: UP000464024).
    Note that ORF1a results in multiple proteins, 
    as shown in figure \ref{fig:covid_genome_and_proteome}.
    Legend: 1 = located in the membrane, 0 = not located in the membrane.
    Topology obtained using 'pureseqtmr' \cite{pureseqtmr}
  }
  \label{fig:covid_topology}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Added PDFs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

These PDFs are added. These will be moved into the article when needed.

### 2020-05-19 covid_dataset_strength/covid_dataset_strength.Rmd
EOD

```{r cache=TRUE}
fasta_filename <- "covid.fasta"
expect_true(file.exists(fasta_filename))
topology <- pureseqtmr::predict_proteome_topology(fasta_filename)
```

@@

```{r}
n_aas <- n_non_tmh_aas + n_tmh_aas
message("proteome size, number of amino acids, n_aas: ", n_aas)
```

@@

>>>>

```{r cache=TRUE}
fasta_filename <- "UP000464024.fasta"
expect_true(file.exists(fasta_filename))
topology <- pureseqtmr::predict_proteome_topology(fasta_filename)
```

@@

```{r}
n_aas <- n_non_tmh_aas + n_tmh_aas
proteome_size <- n_aas
message("number of amino acids, n_aas: ", n_aas)
message("proteome size, proteome_size: ", proteome_size)
```

@@

Get the proteome mutation rate, by converting it from the 
known DNA mutation rate:

```{r}
genome_size <- 30000 # nucleotides

dna_evolution_rate_per_genome_per_year <- 26 # From GISAID
# (Same as bar2020sars, that esimates a DNA 
# evolution rate per nucleotide per year at 0.001)
```

```{r}
rel_size_proteome_to_genome <- 
  proteome_size / genome_size
message("rel_size_proteome_to_genome: ", rel_size_proteome_to_genome)
```

\begin{itemize}
	\item Proteome size: 9701 (from reference proteome)
	\item Size of proteome compared to genome: $\frac{9701}{3 \cdot 10^4} \approx \frac{1}{3}$ \richel{no/little non-coding DNA indeed}
	\item Fraction of mutations that change AA: $\frac{20}{64}$ 
	\item Evolution rate: $\approx 0.00092$ per amino acid per year (as $\frac{1}{3} \cdot \frac{20}{64} \cdot 0.00088$)
	\item TMHs per proteome: 26 (from table \ref{table:covid_tmhs})
	\item AAs per TMHs: 23 (from Bianchi et al.)
	\item AAs per proteome that is TMH: $21 \cdot 23 = 483$
	\item Mutations in TMH per proteome year: $\approx 0.444$ (as $483 \cdot 0.00092$)
	\item Number of sequenced proteomes: 5600 (GISAID, genomes)
	\item Pandemic duration: $\approx \frac{1}{3}$ year
	\item Expected number of observed TMH mutations: $\approx 800$ (as $0.444 \cdot 5600 \cdot \frac{1}{3}$)
	\item Chance that a mutation may lose/acquire epitope binding: $\frac{1}{100}$ \richel{Just a guess, based on weak selection}
	\item Expected number of observed TMH mutations that cause a transition: $\approx 8$ (as $800 \cdot \frac{1}{100}$)
\end{itemize}
